---
title: "RWorksheet_Benliro#3b"
author: "Rica Marie Benliro"
date: "2023-10-15"
output:
  pdf_document: default
  html_document:
  df_print: paged
  word_document: default
---

1. Create a data frame using the table below.

a. Write the codes.

```{r}
household_data <- data.frame(
  respondents = c(1:20),
  sex = c(2,2,1,2,2,2,2,2,2,2,1,2,2,2,2,2,2,2,1,2),
  father_occupation = c(1,3,3,3,1,2,3,1,1,1,3,2,1,3,3,1,3,1,2,1),
  personAtHome = c(5,7,3,8,5,9,6,7,8,4,7,5,4,7,8,8,3,11,7,6),
  siblingAtSchool = c(6,4,4,1,2,1,5,3,1,2,3,2,5,5,2,1,2,5,3,2),
  typeOfHouses = c(1,2,3,1,1,3,3,1,2,3,2,3,2,2,3,3,3,3,3,2)
)

household_data
```

b. Describe the data. Get the structure or the summary of the data
```{r}
summary(household_data)
```

c. Is the mean number of siblings attending is 5?
```{r}
mean(household_data$siblingAtSchool)
```

d. Extract the 1st two rows and then all the columns using the subsetting functions. Write the codes and its output.
```{r}
household_data[1:2, ]
```
e. Extract 3rd and 5th row with 2nd and 4th column. Write the codes and its result.
```{r}
household_data[c(3,5), c(2,4)]
```

f. Select the variable types of houses then store the vector that results as types_houses. Write the codes.

```{r}
types_houses <- household_data$typeOfHouses
types_houses
```

g. Select only all Males respondent that their father occupation was farmer. Write the codes and its output.


```{r}
selection1 <- subset(household_data, sex == 1 & father_occupation == 1)
selection1
```

h. Select only all females respondent that have greater than or equal to 5 number of siblings attending school. Write the codes and its outputs.

```{r}
selection2 <- subset(household_data, sex == 2 & siblingAtSchool >=5)
selection2
```



2. Write a R program to create an empty data frame. Using the following codes:

```{r}
df = data.frame(Ints=integer(),
                Doubles=double(), Characters=character(),
                Logicals=logical(),
                Factors=factor(),
                stringsAsFactors=FALSE)

print("Structure of the empty dataframe:")
print(str(df))
```
a. Describe the results.

The results shows an empty data frame with 5 different data types but no observations since we have not entered any values in the data frame. 


3. Create a .csv file of this. Save it as HouseholdData.csv

```{r}
householdDataCSV <- read.csv(file = "HouseholdData.csv", header = T, stringsAsFactors = F, sep = ",")
householdDataCSV
```

b. Convert the Sex into factor using factor() function and change it into integer.[Legend: Male = 1 and Female = 2]. Write the R codes and its output.
```{r}
householdDataCSV$Sex <- factor(householdDataCSV$Sex, levels = c("Male", "Female"), labels = c(1, 2))
householdDataCSV$Sex <- as.integer(householdDataCSV$Sex)

householdDataCSV
```

c. Convert the Type of Houses into factor and change it into integer. [Legend: Wood = 1; Congrete = 2; Semi-Congrete = 3]. Write the R codes and its output.
```{r}
householdDataCSV$Types.of.Houses <- factor(householdDataCSV$Types.of.Houses, levels = c("Wood", "Congrete", "Semi-congrete"), labels = c(1,2,3))

householdDataCSV$Types.of.Houses <-as.integer(householdDataCSV$Types.of.Houses)

householdDataCSV
```

d. On father’s occupation, factor it as Farmer = 1; Driver = 2; and Others = 3. What is the R code and its output?

```{r}
householdDataCSV$Fathers.Occupation<- factor(householdDataCSV$Fathers.Occupation, levels = c("1","2","3"), labels = c("Farmer", "Driver", "Others"))

householdDataCSV
```

e. Select only all females respondent that has a father whose occupation is driver. Write the codes and its output.

```{r}
female_selections <- subset(householdDataCSV, Sex == 2 & Fathers.Occupation == "Driver")
female_selections
```

f. Select the respondents that have greater than or equal to 5 number of siblings attending school. Write the codes and its output.
```{r}
withSiblings_selection <- subset(householdDataCSV, Siblings.at.School >= 5)
withSiblings_selection
```

4. Interpret the graph.

#In the graph entitled "Sentiments of Twitter Per Day", shows a summary of sentiments scraped from Twitter from July 14, 2020 to July 21, 2020. There are 3 categories of sentiments: Negative, Neutral, and Positive. We can see in the graph that negative sentiments are significantly high in all of the dates. This means that Twitter produces more negative sentiments more than positive which indicates a very toxic and harmful social media environment during the span of the said dates. 

